---
title: "Reproducible Research: Peer Assessment 1"
author: "Jesse Leigh Patsolic"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

First here are the libraries that will be required for this analysis.
```{r libraries}
library(knitr)
library(ggplot2)
```

To begin this analysis we start by loading the activity data provided
for the course and assigning the appropriate classes to each column.
A summary of the data follows:

```{r initialize}
dat <- as.data.frame(read.csv("./activity.csv"))

dat$steps <- as.numeric(dat$steps)
dat$date <- as.Date(strptime(dat$date, format="%Y-%m-%d"))
dat$interval <- as.numeric(dat$interval)

summary(dat)
```


## What is mean total number of steps taken per day?

There are 61 days of measurments at 5-minute intervals.  
I begin by creating a data frame with one row for each unique 
day of measurements.  Then I add an empty column `days$sum` 
to hold the daily sums as they are calculated in the 
following `for` loop. 


```{r days1}
days <- data.frame(day = as.Date(unique(dat$date)), sum = NA)

for(i in 1:dim(days)[1]){
    days$sum[i] <- sum(dat[dat$date == days[i,1],"steps"],na.rm=T)
    }

days$sum <- as.numeric(days$sum)
```

Now we have a data frame with totals for each day.  We can now plot the
histogram.  Even though the `barplot` function is used the bin widths
are length 1 and therefore the heights give the percantage of steps for
each day. 

```{r plots1}
png("./figure/hist01.png", height=480, width=480)
barplot(days$sum, col="blue")
dev.off()

barplot(days$sum, col="blue")
```

The mean and median of total number of steps taken each day are
presented below.

```{r meanmedian}
mean(days$sum)
median(days$sum)
```

## What is the average daily activity pattern?

The following will create a time series plot of the averages for each of
the 288 intervals measured each day.

```{r plots2}
ints <- data.frame(intervals = unique(dat$int))
ints$avg <- NA
for(i in 1:length(ints$int)){
    ints$avg[i] <- 
        sum(dat[dat$int == ints$int[i],"steps"], na.rm=TRUE)/61;
    }

intMax <- ints[ints$avg == max(ints$avg),]

plot(ints$avg~ints$int, 
     type="l", 
     col=3, 
     lwd=2,
     xlab="Intervals", 
     ylab="Average Steps")
title("Interval Averages") 
points(ints[ints$avg == max(ints$avg),])
text(ints[ints$avg == max(ints$avg),],pos=4, 
        label=paste(intMax[,1],round(intMax[,2],2),sep=","))
```

The 835th interval has the maximum average of `r round(intMax[,2],2)`.


## Imputing missing values

```{r NAs}
NAs <- length(which(!complete.cases(dat)))
```

There are `r NAs` rows with missing values in the data. 
The original data will be copied into `dat2` and the missing values 
will be imputed with the mean for that interval as computed in the
graph above.

```{r imputing}
dat2 <- dat

for(i in 1:dim(dat2)[1]){
    if(is.na(dat2[i,"steps"])){
        dat2[i,1] <- ints[ints$int == dat2[i,3],2]
        }
    }
```

Now `dat2` has been imputed and we may continue with the analysis.
`days2` is constructed as before with the imputed data.


```{r days2}
days2 <- data.frame(day = as.Date(unique(dat2$date)), sum = NA)

for(i in 1:dim(days2)[1]){
    days2$sum[i] <- sum(dat2[dat2$date == days2[i,1],"steps"],na.rm=T)
    }

days2$sum <- as.numeric(days2$sum)
```

```{r plots3}
png("./figure/hist02.png", height=480, width=480)
barplot(days2$sum, col="blue")
dev.off()

barplot(days2$sum, col="blue")
```

To compare the two histograms 

```{r compare}
par(mfrow=c(2,1))

barplot(days$sum, col="blue")
barplot(days2$sum, col="blue")
```


## Are there differences in activity patterns between weekdays and weekends?




