---
title: "Reproducible Research: Peer Assessment 1"
author: "Jesse Leigh Patsolic"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

First here are the libraries that will be required for this analysis.
```{r libraries}
library(ggplot2)
```

To begin this analysis we start by loading the activity data provided
for the course and assigning the appropriate classes to each column.
A summary of the data follows:

```{r initialize}
dat <- as.data.frame(read.csv("./activity.csv"))

dat$steps <- as.numeric(dat$steps)
dat$date <- as.Date(strptime(dat$date, format="%Y-%m-%d"))
dat$interval <- as.numeric(dat$interval)

summary(dat)
```


## What is mean total number of steps taken per day?

There are 61 days of measurments.  I begin by creating a data frame with
one row for each unique day of measurements.  Then I add an empty column
`days$sum` to hold the daily sums as they are calculated in the 
following `for` loop. 



```{r days}
days <- data.frame(day = as.Date(unique(dat$date)), sum = NA)

for(i in 1:dim(days)[1]){
    days$sum[i] <- sum(dat[dat$date == days[i,1],"steps"],na.rm=T)
    }

days$sum <- as.numeric(days$sum)
```

Now we have a data frame with totals for each day.  We can now plot the
histogram.

```{r plots1}

```

The mean and median of total number of steps taken each day are
presented below.

```{r meanmedian}
mean(days$sum)
median(days$sum)
```

## What is the average daily activity pattern?

The following will create a time series plot for each of the 5-minute
intervals with a line indicating the 

```{r plotw2}
plot(dat$int, dat$steps, type="l")

```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
